# **Experimental design** {#Exp_design}

<br><br>

```{r diagram_exp_des, echo=FALSE, fig.cap= "Design workFlow"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = true, width = 4, height = 2]
A [label = 'Hypothesis']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
B1 [label = 'Experimental design']

node [color = black, fillcolor = white, fixedsize = true, width = 6, height = 2]
C1 [label = 'Sampling design \\n and collection']
C2 [label = 'QA/QC']
C3 [label = 'Collection of \\n metadata']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
D1 [label = 'Chemical analysis']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
E1 [label = 'Data analysis']

      
      
# edge statements
edge [color = black]
A -> {B1}
B1 -> {C1 C2 C3}
C1 -> {D1}
C2 -> {C1}
C2 -> {D1}
C3 -> {E1}
D1 -> {E1}

      
      }
      ")


```



## Hypothesis
<br>
xx

## Experimental design
<br>
- Compounds of interest?  
- Amendable to GC and/or LC?  
- Instrumentations available?  

## QA/QC measures
<br>
- Spiking with native before and after extraction to evaluate the extraction efficiency and matrix effect. <br>
- Dilute QC extracts to investigate potential matrix effects(?) <br>
- Field blanks <br>
- Procedural blanks <br>
- Solvent blanks <br>
- Procedural replicates <br>
- Triplicate injections <br>. Should also inject the procedural replicates in triplicates.
- Standard reference material <br>
- Randomization <br>
- Retention time index standard <br>
- Volumetric internal standard added into the final extract to guard against small difference of extract volumes, as well as enable normalization of intensities/areas between samples and batches.Use e.g. 4,4'-Dibromooctafluorobiphenyl which should not be in the samples and is stable. <br>
- 

## Data directory structure  

Before we continue with chemical and data analysis, you should first structure the project folder in your computer where you store all the raw and metadata.Below is a suggestion for file directory structure for your project folder. Try to use underscore(_) in your folder names as space causes some errors in various systems and software.

```{}
YOUR_PROJECT_NAME
│   Project_info.xlsx
│
└───Chemicals_Database
│   │   Chemicals.xlsx
│   │   ...
│   │
│   
└───Literature
│   │   Interesting_paper1.pdf
│   │   ..
│
└───Raw_files
│   │   File_info.xlsx
│   │   ..
│   │   └───LC_positive
│   │   └───LC_negative
│   │   └───GC_orbitrap
│
└───Results_Data_analysis
│   │   Hit_list.xlsx
│   │   Quantification.xlsx
│   │   ..
│   │   └───Peaklist_LC_positive
│   │   └───Peaklist_LC_negative
│
└───Sample_info
│   │   Sample_info.xlsx
│
└───Manuscript
│   │   Manuscript_v1.docx

    
    
    
```


## Sampling design and collection
xx
<br>

What's in a name? <br>
Sample names/ID <br>
Naming sample IDs might seem trivial at first but could help downstream data analysis and also to clarify your sampling design. It is important to have a uniform naming. 
convention and that the sample codes could also specify if a sample is a field blank, procedural blank, replicates, belong to a specific group...
Dont start a sample ID with a number

## Collection of metadata
Metadata is the information about data. In this particular case, it is the relevant information about our samples and chemicals of interest.  
Examples could be the location, group, 





## Chemical analysis
xx
<br>





## Data analysis
xx
<br>

## Workflow examples
<br>


```{r example_workflows, echo=FALSE, fig.cap= "Workflow for nontarget using XX"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = false]
A [label = 'Indoor dust collected \\n among different \\n use categories']

node [color = black, fillcolor = white, fixedsize = false]
B1 [label = 'Collect different groups']
B2 [label = 'QA/QC']


node [color = black, fillcolor = white, fixedsize = false]
C1 [label = 'Sample different groups']
C2 [label = 'Suspect list of Cl/Br compounds']

node [color = black, fillcolor = white, fixedsize = false]
D1 [label = 'Chemical analysis']

node [color = black, fillcolor = white, fixedsize = false]
E1 [label = 'Suspect screening']
E2 [label = 'Nontarget']

F1 [label = 'Mass defect plots']

      
      
# edge statements
edge [color = black]
A -> {B1}
B2 -> {C1}
B1 -> {C1 C2}
C1 -> {D1}
C2 -> {E1}
D1 -> {E1}
E2 -> {F1}

      
      }
      ")


```



```{r example_identification, echo=FALSE, fig.cap= "Workflow for structure elucidation and identification"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = false]
L1 [label = 'Level 1']
L2 [label = 'Level 2']
L3 [label = 'Level 3']
L4 [label = 'Level 4']
L5 [label = 'Level 5']

node [color = black, fillcolor = white, fixedsize = false]
S1 [label = 'Mass defect plot']
S2 [label = 'Molecular networking']
S3 [label = 'Fragmentation prediction']
S4 [label = 'Retention time prediction']
S5 [label = 'HRMS library search']


node [color = black, fillcolor = white, fixedsize = false]
A1 [label = 'Exposure index']
A2 [label = 'USdbdb']
A3 [label = 'Building products database']

      
# edge statements
edge [color = black]


      
      }
      ")


```
