# **Experimental design** {#Exp_design}

<br><br>

```{r diagram_exp_des, echo=FALSE, fig.cap= "Design workFlow"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = true, width = 4, height = 2]
A [label = 'Hypothesis']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
B1 [label = 'Experimental design']

node [color = black, fillcolor = white, fixedsize = true, width = 6, height = 2]
C1 [label = 'Sampling design \\n and collection']
C2 [label = 'QA/QC']
C3 [label = 'Collection of \\n metadata']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
D1 [label = 'Chemical analysis']

node [color = black, fillcolor = white, fixedsize = true, width = 5, height = 2]
E1 [label = 'Data analysis']

      
      
# edge statements
edge [color = black]
A -> {B1}
B1 -> {C1 C2 C3}
C1 -> {D1}
C2 -> {D1}
C3 -> {E1}
D1 -> {E1}

      
      }
      ")


```



## Hypothesis
xx

## Experimental design
<br>
- Compounds of interest?  
- Amendable to GC and/or LC?  
- Instrumentations available?  

## QA/QC
 - Spiking with native before and after extraction to evaluate the extraction efficiency and matrix effect
<br>

## Sampling design and collection
xx
<br>

## Collection of metadata
Metadata is the information about data, in this particular case, the relevant information about our samples and chemicals of interest.  
Examples could be the location, group, 

## Chemical analysis
xx
<br>

## Data analysis
xx
<br>

## Workflow examples
<br>


```{r example_workflows, echo=FALSE, fig.cap= "Workflow for nontarget using XX"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = false]
A [label = 'Indoor dust collected \\n among different \\n use categories']

node [color = black, fillcolor = white, fixedsize = false]
B1 [label = 'Collect different groups']
B2 [label = 'QA/QC']


node [color = black, fillcolor = white, fixedsize = false]
C1 [label = 'Sample different groups']
C2 [label = 'Suspect list of Cl/Br compounds']

node [color = black, fillcolor = white, fixedsize = false]
D1 [label = 'Chemical analysis']

node [color = black, fillcolor = white, fixedsize = false]
E1 [label = 'Suspect screening']
E2 [label = 'Nontarget']

F1 [label = 'Mass defect plots']

      
      
# edge statements
edge [color = black]
A -> {B1}
B2 -> {C1}
B1 -> {C1 C2}
C1 -> {D1}
C2 -> {E1}
D1 -> {E1}
E2 -> {F1}

      
      }
      ")


```




