# **Library search** {#library}

In environmental analysis, NIST MS library is widely used to match a spectrum to a list of standard spectra. However, a major drawback of the NIST MS Search is that now all chemicals are included and all spectral ions are at unit mass (for EI spectra library, also called mainlib)


<br><br>  
## IN SILICO FRAGMENTATION  

There are three different models that have recently been published to prdict the fragmentation patterns of GC-EI MS. 

<br><br>
### MEtfrag 
<br><br>


### Sirius CSI FingerID 
<br><br>



### CFM-ID  
In silico fragmentation:  
Help file: https://sourceforge.net/p/cfm-id/wiki/Home/  

<br>
_Use instructions:_
  
- If you are using EI-MS (GC-MS) data, please use the ei_ms_model provided.  
- Note that lpsolve55.dll must also be included in the same directory as the executables. This file can be found in the development version of LPSolve (e.g. lp_solve_5.5.2.0_dev_win32.zip), which can be downloaded from https://sourceforge.net/projects/lpsolve/files/lpsolve/5.5.2.5/lp_solve_5.5.2.5_dev_win32.zip/download.  
  
Cfm-predict  


COMMAND:  
`cfm-predict.exe <smiles_or_inchi_or_file> <prob_thresh> <param_file> <config_file> <annotate_fragments> <output_file_or_dir> <apply_postproc> <suppress_exceptions>`  
  
<br>
  
EXAMPLE (naphthalene):  
`cfm-predict.exe InChI=1S/C10H8/c1-2-6-10-8-4-3-7-9(10)5-1/h1-8H 0.001 D:/Program/cfm-id-2.4_win32/ei_nn_iso_new/param_output.log D:/Program/cfm-id-2.4_win32/ei_nn_iso_new/param_config.txt 0 D:/Program/cfm-id-2.4_win32/test/output.msp`

<br>

EXAMPLE (.txt file as input):  
`cfm-predict.exe D:/Program/cfm-id-2.4_win32/test/input.txt 0.001 D:/Program/cfm-id-2.4_win32/ei_nn_iso_new/param_output.log D:/Program/cfm-id-2.4_win32/ei_nn_iso_new/param_config.txt 0 D:/Program/cfm-id-2.4_win32/test/output.msp`  



<br><br><br>

## NEIMS  
  
Help files: https://github.com/brain-research/deep-molecular-massspec  

<br><br><br>

## DeepEI  
  
Help files: https://github.com/hcji/DeepEI  
https://github.com/hcji/DeepEI/blob/master/Usage.ipynb  

<br><br><br>


## QCEIMS

RUN IN LINUX
  
  
PREPARATION  
Copy the ".XTBPARAM" folder and the ".mass_raw.agr" file to /home/ORUNET.ORU.SE/twg/ 
  
  
1. Prepare a file with the equilibrium structure of your desired molecule M.
Important: This file has to be named coord and should have the TURBOMOLE coord format.
In most cases you will have an .xyz file. This file can be easily converted by typing:
  
x2t <xyzfile> > coord
  
if you have installed TURBOMOLE. If you do not have TURBOMOLE, you may have to write a
script converting .xyz files to TUROBOMOLE coord files. Be advised that the coord file has
to be in atomic units.
  
CAN THIS BE DONE USING OPENBABEL?
  
2. Prepare an input file called qceims.in. For the input options, see section 4 or the
qceims.in file in the examples folder. If no such file is prepared, default options are:
run GFN1-xTB with 25 times the number of atoms in the molecule trajectories (ntraj).

RUN QCEIMS

set +o noclobbe
  
export PATH=/home/ORUNET.ORU.SE/twg/QCEIMS/:/home/ORUNET.ORU.SE/twg/bin:/home/ORUNET.ORU.SE/twg/.local/bin:/bin:/usr/bin:/opt/thinlinc/bin:/usr/local/bin:/usr/bin/X11:/sbin:/usr/sbin:/usr/local/sbin:/snap/bin:/opt/thinlinc/bin:/opt/SPAdes/SPAdes-3.13.0-Linux/bin:/opt/mauve/mauve_snapshot_2015-02-13:/opt/parsnp/Parsnp-Linux64-v1.2:/opt/prokka/prokka-master/bin:/opt/artemis/artemis
  
cd QCEIMS/example/2-Chloroethanol_GFN2
  
qceims
  
## LIBRARY SEARCH  
xx

```{r readMGF, eval=FALSE, include=FALSE, echo=TRUE}
# from https://github.com/daniellyz/MergeION/blob/master/R/readMGF2.R

# Extracting metadata and a list of spectra from mgf file
# The function extracts a dataframe of metadata and a list of spectra from mgf file or


readMGF2<-function(con){


  options(stringsAsFactors = FALSE)
  options(warn=-1)

  # if (is.character(con)){
  #   if (file_ext(con)!="mgf"){
  #     stop("The file extension of your input library must be mgf!")
  #   }} else {
  #     stop("The input must be the name of the mgf file!")}

  #####################################
  ### Reading from spectral library:
  #####################################

  db = readMgfData(con, verbose = FALSE)
  metadata = fData(db)
  N = nrow(metadata)

  # From a MSnBase object to a list of spectra m/z intensity
  spectrum_list=list()
  for (i in 1:N){
    spectrum_list[[i]]=cbind(db[[i]]@mz,db[[i]]@intensity)
  }

  ####################
  ### Return results:
  ####################

  library = list()
  if (N>0){
    library$sp = spectrum_list
    library$metadata = metadata
  }
  return(library)
}


test <- readMGF2("D:\\Raw_data\\Dust_Florian\\GC\\test\\test.mgf")
unknown1 <- test$sp[[1]]



```

  
  
### LIB2NIST  
Convert your library list into NIST library format using LIB2NIST command line (in order to preserve the accurate mass). See command line help file for arguments.  
  
COMMAND:  
`lib2nist.exe /log9 Mylib.log /OutLib /StdRounding:N /MsmsOnly:Y /AccuratePeakMZ /PrecurMzDecPlaces=keep /PeakMzDecPlaces=keep /UseSubset:N <path to msp file> <output path> =<new name of library>`  
    
EXAMPLES:  
`lib2nist64.exe /log9 Mylib.log /OutLib /StdRounding:N /MsmsOnly:Y /AccuratePeakMZ /PrecurMzDecPlaces=keep /PeakMzDecPlaces=keep /UseSubset:N D:\Projects\Suspect_lists\Spectral_databases\RECETOX_GC-EI_MS_20201028.msp D:\Program\NIST14\ =RECETOX_GC-EI_MS_20201028`  
  
<br><br><br>

### MSPepSearch  
    
Use MSPepSearch to find similar spectra  
  
  
EXAMPLES:  
`MSPepSearch64.exe Gusviqh /ZI 0.1 /ZIPPM 20 /MPPM 30 /MzLimits 50 -1 /MinMF 10 /OnlyFound /HITS 5 /LIB D:\Raw_data\Dust_Florian\GC\test\Mylib /INP D:\Raw_data\Dust_Florian\GC\test\input.msp /OUTMGF D:\Raw_data\Dust_Florian\GC\test\test.mgf /OUTTAB D:\Raw_data\Dust_Florian\GC\test\test.tsv /OutMW`
  
  
`MSPepSearch64.exe Gusviqh /ZI 0.1 /ZIPPM 20 /MPPM 30 /MzLimits 50 -1 /MinMF 100 /OnlyFound /HITS 5 /LIB C:\NIST14\NIST_contaminants_orbitrap /INP D:\Projects\Mexico_Air\NIST_Mexico.msp /OUTMGF D:\Projects\Mexico_Air\test\test.mgf /OUTTAB D:\Projects\Mexico_Air\test\test.tsv`  
  
  
`MSPepSearch64.exe Gusviqh /ZI 0.01 /ZIPPM 10 /MPPM 10 /MzLimits 50 -1 /MinMF 500 /OnlyFound /HITS 5 /LIB C:\NIST14\LCMs_GC_Orbitrap /INP D:\Raw_data\Dust_Florian\GC\test\11.msp /OUTMGF D:\Raw_data\Dust_Florian\GC\test\test.mgf /OUTTAB D:\Raw_data\Dust_Florian\GC\test\test.tsv`  



```{r eval=FALSE, include=FALSE, echo=TRUE}


system("D:\\Program\\MSPepSearch\\MSPepSearch64.exe Gusviqh /ZI 0.01 /ZIPPM 10 /MPPM 10 /MzLimits 50 -1 /MinMF 500 /OnlyFound /HITS 5 /LIB C:\\NIST14\\LCMs_GC_Orbitrap /INP D:\\Raw_data\\Dust_Florian\\GC\\test\\11.msp /OUTMGF D:\\Raw_data\\Dust_Florian\\GC\\test\\test.msp /OUTTAB D:\\Raw_data\\Dust_Florian\\GC\\test\\test.tsv")


# not working yet!
test <- run(command = "D:/Program/MSPepSearch/MSPepSearch64.exe", args = "Gusviqh /INP D:\\Raw_data\\Dust_Florian\\GC\\test\\11.msp")
cat(test$stderr)

hits <- read_tsv("D:\\Raw_data\\Dust_Florian\\GC\\test\\test.tsv", skip = 3)
hits <- hits %>% 
  slice_head(n = length(hits[[1]])-2)

spectra <- readMgfData(filename = "D:\\Raw_data\\Dust_Florian\\GC\\test\\test.mgf")

plot(spectra[[1]], spectra[[2]])


```

