# Example workflow: indoor dust {#example_dust}

A walkthrough for indoor dust analysis

<br><br>  




```{r example_workflows_dust, echo=FALSE, fig.cap= "Workflow for nontarget using XX"}

# Diagrammer: rich-iannone.github.io/DiagrammeR/graphviz_and_mermaid.html

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false, rankdir = LR]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = false]
# A [label = 'Indoor dust collected \\n among different \\n use categories']
# 
# node [color = black, fillcolor = white, fixedsize = false]
# B1 [label = 'Collect different groups']
# B2 [label = 'QA/QC']
# 
# 
# node [color = black, fillcolor = white, fixedsize = false]
# C1 [label = 'Sample different groups']
# C2 [label = 'Suspect list of Cl/Br compounds']

node [color = black, fillcolor = white, fixedsize = false]
D1 [label = 'Sample extraction']
D1_1 [label = 'Hexane 3 mL']
D1_2 [label = 'Hexane: Acetone (1:1) 3 mL']
D1_3 [label = 'Toluene 3 mL']
D2 [label = 'Combine and split 1:1']
D3_1 [label = 'GC']
D3_2 [label = 'LC']
D4 [label = 'split 1:1']



E1 [label = 'GC-HRMS']
E2 [label = 'LC-HRMS']
# 
# 
# node [color = black, fillcolor = white, fixedsize = false]
# F1 [label = 'Suspect screening']
# F2 [label = 'Nontarget']
# 
# G1 [label = 'Mass defect plots']

      
      
# edge statements
edge [color = black]
# A -> {B1}
# B2 -> {C1}
# B1 -> {C1 C2}
# C1 -> {D1}

D1 -> {D1_1}
D1_1 -> {D1_2}
D1_2 -> {D1_3}
D1_1 -> {D2}
D1_2 -> {D2}
D2 -> {D3_1}
D2 -> {D3_2}
D3_1 -> {D4}
D1_3 -> {D4}

D3_2 -> {E2}
D4 -> {E1}
D1_3 -> {E1}
      
      }
      ")


```


```{r data processing, echo=FALSE}

grViz("
digraph Scheme {
      
# graph statement
graph [layout = dot, overlap = false]

# node statements
node [shape = box,
fontname = Helvetica, fontsize = 35, style = filled]
      
node [color = black, fillcolor = white, fixedsize = false]

ZERO [label = 'Deconvolution']

A [label = 'Target analysis']
B [label = 'Suspect list screening']

C [label = 'NIST library screening']
C1 [label = 'Molecular ion confirmation']
C2 [label = 'Retention time index']


D [label = 'Mass defect \\n (selected mz, rt and int of annotated compounds']

subgraph cluster2 {
E [label = 'Annotation']
ID1 [label = 'Confirm EIC peaks']
}

F [label = 'Molecular networking']
G [label = 'Multivariate statistics']

ID2 [label = 'Peak/Component list with \\n partial annotation']

Q1 [label = 'Semiquantify across samples']      
      
# edge statements

edge [color = black]

ZERO -> {A B C}
A -> {E}
B -> {E}

{C1 C2} -> {C}
C -> {E}

E -> {ID1}
ID1 -> {ID2}
ID2 -> {F}
ID2 -> {G}
ID2 -> {D}
ID2 -> {Q1}

}
")


```


<br>

## **GC orbitrap HRMS workflow**
This code uses XCMS with MF -> CAMERA preprocessing used in : Stettin, D.; Poulin, R.X.; Pohnert, G. Metabolomics Benefits from Orbitrap GC–MS—Comparison of Low- and High-Resolution GC–MS. Metabolites 2020, 10, 143. See the supplementary material for original code. The code was modified here to also include retention time index and also remove zero intensity peaks after peak grouping. Additional metainformation for the msp file can also be added in the "extra" object.


<br>
```{r XCMS MF, echo=TRUE, eval=FALSE}


library(processx)
library(xcms)
library(CAMERA)
library(metaMS)
library(dplyr)
library(stringr)

#first define the working directory (folder with experimental group folders inside)
# mzMLfiles <- list.files(path = "/home/ORUNET.ORU.SE/twg/Dioxinlab/GC-Orbitrap data/Florian/Dust NTA 20200513/mzXML/",pattern = ".mzXML", recursive = TRUE, full.names = TRUE)
# Remove files containing "Hex", blank samples
# mzMLfiles <- mzMLfiles[-grep("Hex", mzMLfiles)]

sample_info <- readxl::read_xlsx("/home/ORUNET.ORU.SE/twg/Windows_home/Dust_Florian/Sample_list_NTA_dust.xlsx",
                               sheet = "Sample")
# filter only samples
sample_info <- sample_info %>%
  dplyr::filter(!(group %in% c("blank_solvent")))

mzXMLfiles <- as.character(sample_info$filename)

pd <- data.frame(sample_name = paste0(sample_info$new_codes, ".mzXML"),
                 class = sample_info$group,
                 stringsAsFactors = FALSE)
# pd <- pd %>% dplyr::filter(sample_group != "IS") %>% dplyr::filter(sample_group != "RTI")


min.clustersize <- 17 #how many fragments need to be in a group for it to be included in the results. 5 - trace compounds; 20 - mostly high quality spectra
useOriginalCode(TRUE) #otherwise matchedFilter will allocate a huge amount of RAM when using step = 0.001

#Refer to XCMS and CAMERA documentation for information about parameters and functions
#These are the parameters used for high rez GC-Orbitrap data
xs1 <- xcmsSet(mzXMLfiles, 
               phenoData = pd,
               method = "matchedFilter", 
               fwhm = 3, step = 0.001, 
               steps = 2, max = 1000, 
               snthresh = 3, mzdiff = 0.002)

xs2 <- group(xs1, method = "density", bw = 5, mzwid = 0.002, minsamp = 1, minfrac = 0, max = 1000)

xs3 <- retcor(xs2, method = "obiwarp")

xs4 <- group(xs3, method = "density", bw = 2, mzwid = 0.6, minsamp = 1, minfrac = 0, max = 1000)

xs4 <- group(xs3, method = "density", bw = 2, mzwid = 0.002, minsamp = 1, minfrac = 0, max = 1000)

xs5 <- fillPeaks(xs4)

xs6 <- xsAnnotate(xs = xs5, sample=NA, polarity = "positive")

xs7 <- groupFWHM(xs6, sigma = 6 , perfwhm = 2, intval = "maxo")

xs8 <- groupCorr(xs7, cor_eic_th = 0.8, calcCaS = TRUE)


peaktable <- getPeaklist(xs8, intval="into")
write.csv(peaktable, file = "untreated_peaktable_MatchedFilter.csv") #This creates a peaktable .csv file with all m/z features sorted into "compounds"

#the following script identifies all "compounds" considered too small according to min.clustersize
#you don't need to modify anything, that run the next part as is
peaktable <- getPeaklist(xs8, intval="maxo")
pcgroups <- sort(unique(as.numeric(peaktable$pcgroup)))
lspectra <- NULL
extra <- NULL
small.clusters <- NULL
big.clusters <- NULL
n <- 1
for(x in pcgroups) {
  clustersize <- length(peaktable[peaktable$pcgroup==x,ncol(peaktable)])
  if(clustersize < as.numeric(min.clustersize))
  {
    small.clusters <- c(small.clusters, x)
  }
  else
  {
    big.clusters <- c(big.clusters, x)
    gruppe1 <- peaktable[peaktable$pcgroup==x,]
    gruppe1 <- gruppe1[, -(2:(ncol(gruppe1)-3-(length(mzXMLfiles))))]
    gruppe1 <- gruppe1[, -(ncol(gruppe1):(ncol(gruppe1)-2))]
    decider <- NULL
    for(i in 2:ncol(gruppe1))
    {
      decider <- c(decider, sum(gruppe1[, i]))
    }
    highest <- which(decider==max(decider), arr.ind = TRUE)
    gruppe1 <- data.frame(gruppe1[, 1], gruppe1[, highest[1]+1])
    colnames(gruppe1) <- c("mz", "int")
    lspectra[[n]] <- gruppe1
    n <- n+1
  }
}

# remove peaks with zero intensities

lspectra <- lapply(lspectra, function(x) {x[x$int != 0,]})


reduced.peaktable <- getReducedPeaklist(xs8, method = "sum", intval = "into", default.adduct.info = "maxint", mzrt.range = FALSE, npeaks.sum = FALSE, cleanup = FALSE)
if(is.null(small.clusters)==FALSE) {
  for(z in small.clusters)
  {
    reduced.peaktable <- reduced.peaktable[-(which(reduced.peaktable[, ncol(reduced.peaktable)]==z)), ]
  }
}
write.csv(reduced.peaktable, file = "peaktable_MatchedFilter.csv") #This creates a peaktable .csv file with only "compounds" instead of m/z features
extra <- data.frame(Name = paste("Unknown", big.clusters, "RT =",round(reduced.peaktable$rt/60, 2) ), Class = "Unknown", RT = round(reduced.peaktable$rt/60, 2),  stringsAsFactors = FALSE)

####KOVATS#####
#Adding Kovats retention index to the extra obejct to write to msp
data <- data.frame(rt = extra$RT)
alkaneSeries <- data.frame(Num = c(11, 13, 15, 17, 19, 21, 23, 25),
                           rt = c(4.90, 8.00, 11.13, 14.05, 16.70, 19.37, 22.45, 25.77))

RI <- vector(length = nrow(data))
for (i in seq_len(nrow(data))) {
  m <- dplyr::case_when(
  data$rt[i] >= alkaneSeries$rt[1] & data$rt[i] < alkaneSeries$rt[2] ~ alkaneSeries$Num[1],
  data$rt[i] >= alkaneSeries$rt[2] & data$rt[i] < alkaneSeries$rt[3] ~ alkaneSeries$Num[2],
  data$rt[i] >= alkaneSeries$rt[3] & data$rt[i] < alkaneSeries$rt[4] ~ alkaneSeries$Num[3],
  data$rt[i] >= alkaneSeries$rt[4] & data$rt[i] < alkaneSeries$rt[5] ~ alkaneSeries$Num[4],
  data$rt[i] >= alkaneSeries$rt[5] & data$rt[i] < alkaneSeries$rt[6] ~ alkaneSeries$Num[5],
  data$rt[i] >= alkaneSeries$rt[6] & data$rt[i] < alkaneSeries$rt[7] ~ alkaneSeries$Num[6],
  data$rt[i] >= alkaneSeries$rt[7] & data$rt[i] <= alkaneSeries$rt[8] ~ alkaneSeries$Num[7]
)

n <- dplyr::case_when(
  data$rt[i] >= alkaneSeries$rt[1] & data$rt[i] < alkaneSeries$rt[2] ~ alkaneSeries$Num[2],
  data$rt[i] >= alkaneSeries$rt[2] & data$rt[i] < alkaneSeries$rt[3] ~ alkaneSeries$Num[3],
  data$rt[i] >= alkaneSeries$rt[3] & data$rt[i] < alkaneSeries$rt[4] ~ alkaneSeries$Num[4],
  data$rt[i] >= alkaneSeries$rt[4] & data$rt[i] < alkaneSeries$rt[5] ~ alkaneSeries$Num[5],
  data$rt[i] >= alkaneSeries$rt[5] & data$rt[i] < alkaneSeries$rt[6] ~ alkaneSeries$Num[6],
  data$rt[i] >= alkaneSeries$rt[6] & data$rt[i] < alkaneSeries$rt[7] ~ alkaneSeries$Num[7],
  data$rt[i] >= alkaneSeries$rt[7] & data$rt[i] <= alkaneSeries$rt[8] ~ alkaneSeries$Num[8]
)


RI[i] <- round(100*n + 100*(m-n) * (data$rt[i] - alkaneSeries[alkaneSeries$Num == n,]$rt)/(alkaneSeries[alkaneSeries$Num == m,]$rt - alkaneSeries[alkaneSeries$Num == n,]$rt), 0)
}

extra <- cbind(extra, RI)

######END KOVATS######

# create the msp object with spectra and all meta information in the extra object (NAME, RETENTIONTIME, RETENTIONINDEX)
export.msp <- construct.msp(lspectra, extra)

write.msp(export.msp, file = "spectra_MatchedFilter20201030.msp") #This creates a NIST MS Search compatible .msp file with all compound pseudospectra



```


This code uses XCMS with centWave. Check differences and skip first steps to combine both workflows

```{r XCMS CW, echo=TRUE, eval=FALSE}

#first define the working directory (folder with experimental group folders inside)
# mzMLfiles <- list.files(path = "/mzXML/",pattern = ".mzXML", recursive = TRUE, full.names = TRUE)
# Remove files containing "Hex", blank samples
# mzMLfiles <- mzMLfiles[-grep("Hex", mzMLfiles)]

sample_info <- readxl::read_xlsx("Sample_list_NTA_dust.xlsx",
                               sheet = "Sample_xcms")
# filter only samples
sample_info <- sample_info %>%
  dplyr::filter(!(group %in% c("blank_solvent")))

pd <- data.frame(sample_name = paste0(sample_info$analysis, ".mzXML"),
                 sample_group = sample_info$group,
                 stringsAsFactors = FALSE)

mzXMLfiles <- paste0(sample_info$path, "/", sample_info$analysis, ".mzXML")



# Use subset to test
pd <- pd[1:2,]
mzXMLfiles <- mzXMLfiles[1:2]

raw_data <- readMSData(mzXMLfiles, 
                         pdata = new("NAnnotatedDataFrame", pd),
                         mode = "onDisk")

# pd <- pd %>% dplyr::filter(sample_group != "IS") %>% dplyr::filter(sample_group != "RTI")

#-----Find features----#

# centWave params
cwp <- CentWaveParam(ppm = 5,
                     peakwidth = c(3, 45),
                     snthresh = 10,
                     prefilter = c(1,30000),
                     mzCenterFun = "wMean",
                     integrate = 1L,
                     mzdiff = -0.001,
                     fitgauss = FALSE,
                     noise = 30000,
                     verboseColumns = FALSE)


xs1 <- findChromPeaks(mzXMLfiles, param = cwp)

# use model peak to evaluate process
rtr <- c(1440, 1460)
mzr <- c(340.238, 340.242)
chr_raw <- chromatogram(mzXMLfiles, mz = mzr, rt = rtr)
chr_mzr <- chromatogram(xs1, mz = mzr, rt = rtr)
group_colors <- paste0(brewer.pal(3, "Set1")[1:2], "60")
sample_colors <- group_colors[xs1$sample_group]


#----Group features 1----#

pdp <- PeakDensityParam(
  sampleGroups = pd$sample_group,
  bw = 10,
  minFraction = 0,
  minSamples = 1,
  binSize = 0.002,
  maxFeatures = 1000)


xs2 <- groupChromPeaks(xs1, param = pdp)


plotChromPeakDensity(chr_mzr, col = sample_colors, param = pdp)


#----retention time correction----#

obip <- ObiwarpParam(binSize = 0.05,  # need to check optimal binSize
                     centerSample = integer(),
                     response = 1L,
                     distFun = "cor_opt",
                     gapInit = numeric(),
                     gapExtend = numeric(),
                     factorDiag = 2,
                     factorGap = 1,
                     localAlignment = FALSE,
                     initPenalty = 0,
                     subset = integer(),
                     subsetAdjust = c("average", "previous"))

xs3 <- adjustRtime(xs2, param = obip)

# Get the base peak chromatograms
bpis_adj <- chromatogram(xs3, aggregationFun = "max", include = "none")
par(mfrow = c(2, 1), mar = c(4.5, 4.2, 1, 0.5))
plot(bpis_adj)
# Plot also the difference of adjusted to raw retention time.
plotAdjustedRtime(xs3)

par(mfrow = c(2, 1))
## Plot the raw data
plot(chr_raw)

## Extract the chromatogram from the adjusted object
chr_adj <- chromatogram(xs3, rt = rtr, mz = mzr)
plot(chr_adj, peakType = "none")



#----Group features 2----#
# After retention time correction, the rt values are modified and additional grouping is needed.

pdp2 <- PeakDensityParam(
  sampleGroups = pd$sample_group,
  bw = 5,
  minFraction = 0,
  minSamples = 1,
  binSize = 0.2,
  maxFeatures = 1000)


xs4 <- groupChromPeaks(xs3, param = pdp2)

xs4_chrom <- chromatogram(xs4, mz = mzr, rt = rtr)
plotChromPeakDensity(xs4_chrom, col = sample_colors, param = pdp2)

# iteratively decrease the binSize
pdp3 <- PeakDensityParam(
  sampleGroups = pd$sample_group,
  bw = 2,
  minFraction = 0,
  minSamples = 1,
  binSize = 0.002,
  maxFeatures = 1000)

xs4 <- groupChromPeaks(xs4, param = pdp3)

xs4_chrom <- chromatogram(xs4, mz = mzr, rt = rtr)
plotChromPeakDensity(xs4_chrom, col = sample_colors, param = pdp3)

#----Fill missing features----#


xs5 <- fillChromPeaks(xs4)

## Check missing values before filling in peaks
apply(featureValues(xs5, filled = FALSE), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))

## Missing values after filling in peaks
apply(featureValues(xs5), MARGIN = 2,
      FUN = function(z) sum(is.na(z)))



#---- QC ----#

# Extract chromatograms of 4 features
feature_chroms <- featureChromatograms(xs5, features = 10000:10004)

plot(feature_chroms)


boxplot(featureValues(xs5, value="into") +1, 
        #col=as.numeric(sampclass(mtbls2Set))+1, 
        log="y", las=2)

sdThresh <- 4.0 ## Filter low-standard deviation rows for plot
data <- log(featureValues(xs5))+1
pca.result <- pca(data, nPcs=3)
plotPcs(pca.result, type="loadings", 
        #col=as.numeric(sampclass(mtbls2Set))+1
        )




# It is possible to use the retention time correction and grouping step in an iterative way if needed. 
# Once you perform your last adjustRtime step and thus your last grouping step, you will obtain your final peak list (i.e. final list of ions)


#----Annotation using CAMERA----#

# Since CAMERA has not yet been ported to XCMSnExp,we need to convert to xcmsSet. 
# Note that the conversion only makes sense for somple XCMSnSets, 
# without e.g. MS level filtering (where CAMERA would then extract the wrong peaks)

xs6 <- as(xs5, "xcmsSet")

xs6 <- xsAnnotate(xs = xs6, sample=NA, polarity = "positive")

xs7 <- groupFWHM(xs6, sigma = 6 , perfwhm = 2, intval = "maxo")

xs8 <- groupCorr(xs7, cor_eic_th = 0.8, calcCaS = TRUE)


#how many fragments need to be in a group for it to be included in the results. 5 - trace compounds; 20 - mostly high quality spectra
min.clustersize <- 20 

peaktable <- getPeaklist(xs8, intval="into")
write.csv(peaktable, file = "test_peaktable.csv") #This creates a peaktable .csv file with all m/z features sorted into "compounds"

#the following script identifies all "compounds" considered too small according to min.clustersize
peaktable <- getPeaklist(xs8, intval="maxo")

# Replaces NA in intensities with zero
peaktable <- peaktable %>%
  dplyr::mutate(dplyr::across(dplyr::starts_with("X"), ~tidyr::replace_na(., 0)))


pcgroups <- sort(unique(as.numeric(peaktable$pcgroup)))
lspectra <- NULL
extra <- NULL
small.clusters <- NULL
big.clusters <- NULL
n <- 1
for(x in seq_along(pcgroups)) {
  clustersize <- length(peaktable[peaktable$pcgroup==x,ncol(peaktable)])
  if(clustersize < as.numeric(min.clustersize))
  {
    small.clusters <- c(small.clusters, x)
  }
  else
  {
    big.clusters <- c(big.clusters, x)
    gruppe1 <- peaktable[peaktable$pcgroup==x,]
    gruppe1 <- gruppe1[, -(2:(ncol(gruppe1)-3-(length(mzXMLfiles))))]
    gruppe1 <- gruppe1[, -(ncol(gruppe1):(ncol(gruppe1)-2))]
    decider <- NULL
    
    for(i in 2:ncol(gruppe1))
    {
      decider <- c(decider, sum(gruppe1[, i]))
    }
    highest <- which(decider==max(decider), arr.ind = TRUE)
    gruppe1 <- data.frame(gruppe1[, 1], gruppe1[, highest[1]+1])
    colnames(gruppe1) <- c("mz", "int")
    lspectra[[n]] <- gruppe1
    n <- n+1
  }
}


# remove peaks with zero intensities
lspectra <- lapply(lspectra, function(x) {x[x$int != 0,]})

reduced.peaktable <- getReducedPeaklist(xs8, method = "sum", intval = "into", default.adduct.info = "maxint", mzrt.range = FALSE, npeaks.sum = FALSE, cleanup = FALSE)
if(is.null(small.clusters)==FALSE) {
  for(z in small.clusters)
  {
    reduced.peaktable <- reduced.peaktable[-(which(reduced.peaktable[, ncol(reduced.peaktable)]==z)), ]
  }
}
write.csv(reduced.peaktable, file = "test_peaktable.csv") #This creates a peaktable .csv file with only "compounds" instead of m/z features

extra <- data.frame(Name = paste("Unknown", big.clusters, "RT =",round(reduced.peaktable$rt/60, 2) ), Class = "Unknown", RT = round(reduced.peaktable$rt/60, 2),  stringsAsFactors = FALSE)

####KOVATS#####
#Adding Kovats retention index to the extra obejct to write to msp
data <- data.frame(rt = extra$RT)
alkaneSeries <- data.frame(Num = c(11, 13, 15, 17, 19, 21, 23, 25),
                           rt = c(4.90, 8.00, 11.13, 14.05, 16.70, 19.37, 22.45, 25.77))

RI <- vector(length = nrow(data))
for (i in seq_len(nrow(data))) {
  m <- dplyr::case_when(
  data$rt[i] >= alkaneSeries$rt[1] & data$rt[i] < alkaneSeries$rt[2] ~ alkaneSeries$Num[1],
  data$rt[i] >= alkaneSeries$rt[2] & data$rt[i] < alkaneSeries$rt[3] ~ alkaneSeries$Num[2],
  data$rt[i] >= alkaneSeries$rt[3] & data$rt[i] < alkaneSeries$rt[4] ~ alkaneSeries$Num[3],
  data$rt[i] >= alkaneSeries$rt[4] & data$rt[i] < alkaneSeries$rt[5] ~ alkaneSeries$Num[4],
  data$rt[i] >= alkaneSeries$rt[5] & data$rt[i] < alkaneSeries$rt[6] ~ alkaneSeries$Num[5],
  data$rt[i] >= alkaneSeries$rt[6] & data$rt[i] < alkaneSeries$rt[7] ~ alkaneSeries$Num[6],
  data$rt[i] >= alkaneSeries$rt[7] & data$rt[i] <= alkaneSeries$rt[8] ~ alkaneSeries$Num[7]
)

n <- dplyr::case_when(
  data$rt[i] >= alkaneSeries$rt[1] & data$rt[i] < alkaneSeries$rt[2] ~ alkaneSeries$Num[2],
  data$rt[i] >= alkaneSeries$rt[2] & data$rt[i] < alkaneSeries$rt[3] ~ alkaneSeries$Num[3],
  data$rt[i] >= alkaneSeries$rt[3] & data$rt[i] < alkaneSeries$rt[4] ~ alkaneSeries$Num[4],
  data$rt[i] >= alkaneSeries$rt[4] & data$rt[i] < alkaneSeries$rt[5] ~ alkaneSeries$Num[5],
  data$rt[i] >= alkaneSeries$rt[5] & data$rt[i] < alkaneSeries$rt[6] ~ alkaneSeries$Num[6],
  data$rt[i] >= alkaneSeries$rt[6] & data$rt[i] < alkaneSeries$rt[7] ~ alkaneSeries$Num[7],
  data$rt[i] >= alkaneSeries$rt[7] & data$rt[i] <= alkaneSeries$rt[8] ~ alkaneSeries$Num[8]
)


RI[i] <- round(100*n + 100*(m-n) * (data$rt[i] - alkaneSeries[alkaneSeries$Num == n,]$rt)/(alkaneSeries[alkaneSeries$Num == m,]$rt - alkaneSeries[alkaneSeries$Num == n,]$rt), 0)
}

extra <- cbind(extra, RI)

######END KOVATS######

# create the msp object with spectra and all meta information in the extra object (NAME, RETENTIONTIME, RETENTIONINDEX)
export.msp <- construct.msp(lspectra, extra)

write.msp(export.msp, file = "test_spectra.msp") #This creates a NIST MS Search compatible .msp file with all compound pseudospectra


```




## **LC-HRMS using Waters MSe DIA workflow**


```{r startup}
library(xcms)
library(magrittr)
library(RAMClustR)
```

# Workflow for centroiding of raw file, peak picking and correlation of MS1 and MS2


## First convert unifi data using MSConvert

Options: 
output format: mzML
binary encoding precision: 64-bit
Write index: check
TPP compatibility: check
use zlib compression: uncheck
package in gzip: uncheck


Filters:
1. TitleMaker
2. msLevel: 1-2  (need to click "Add" to add filter)

## Centroiding raw data from mzML using MSnbase
```{r}

library(xcms)
library(magrittr)

# mzML file was created by MSConvert without centroiding but only function 1 and function 2 (MS channel 1 and 2).
# Lock mass channel (Channel 3) is therefore removed by MSConvert

# A1. Smoothing and then centroiding of mzML raw data file

data_sg_cent <- data_prof %>%
  smooth(method = "SavitzkyGolay", halfWindowSize = 4L) %>%
  pickPeaks()

## A2. Write the centroided data to files with the same names in the current directory

writeMSData(data_sg_cent, file = "D:/Raw_data/Test_centroidingMSnbase/STD1_1_A,3_1_cent.mzML")

```



```{r}

## B1. Read the centroided data.

files <- "D:/Raw_data/Test_centroidingMSnbase/STD1_1_A,3_1_cent.mzML"

sample.names <- data.frame(sample_name = c("STD1"),
                           sample_group = c("STD"))


raw_data <- readMSData(files = files, 
                         pdata = new("NAnnotatedDataFrame", sample.names),
                         mode = "onDisk")

raw_data_info <- dplyr::glimpse(fData(raw_data))


## B2. Peak detection for both MS1 and MS2

# Set parameters specific for MS instrument (below is for Waters G2 XS qtof)
ms.res <- 40000
peak.width <- c(10,60)
ms.fwhm <- round(550/ms.res, digits = 3)
ppm <- round(1.5*(1000000*ms.fwhm)/(550*2.355))
sn.thresh <- 10 


# Peak picking for MS1
cwp1 <- CentWaveParam(
  ppm = ppm,
  peakwidth = peak.width,
  snthresh = sn.thresh,
  prefilter = c(3, 5000),
  mzCenterFun = "wMean",
  integrate = 2L,
  mzdiff = -0.001,
  fitgauss = TRUE,
  noise = 2000,
  verboseColumns = TRUE,
  roiList = list(),
  firstBaselineCheck = TRUE,
  roiScales = numeric(),
  extendLengthMSW = FALSE
)

xdata <- findChromPeaks(raw_data, param = cwp1, msLevel = 1L)


# Peak picking for MS2
cwp2 <- CentWaveParam(
  ppm = ppm,
  peakwidth = peak.width,
  snthresh = sn.thresh,
  prefilter = c(3, 100),
  mzCenterFun = "wMean",
  integrate = 2L,
  mzdiff = -0.001,
  fitgauss = TRUE,
  noise = 100,
  verboseColumns = TRUE,
  roiList = list(),
  firstBaselineCheck = TRUE,
  roiScales = numeric(),
  extendLengthMSW = FALSE
)

xdata <- findChromPeaks(xdata, param = cwp2, msLevel = 2L, add = TRUE)

# Check that MS1 and MS2 has been peak picked
table(chromPeakData(xdata)$ms_level)

isolationWindowTargetMz(xdata)
isolationWindow(xdata)
chromPeakData(xdata)

# B4. Find and correlate MS2 peaks to MS1
xdata <- findChromPeaksIsolationWindow(xdata, param = cwp2)

xspectra <- reconstructChromPeakSpectra(xdata,
                                        expandRt = 0,
                                        diffRt = 2,
                                        minCor = 0.8,
                                        intensity = "maxo",
                                        peakId = rownames(chromPeaks(xdata, msLevel = 1L)),
                                        BPPARAM = bpparam())



xspectra

## Manually check MS2 spectra that "belongs" to a precursor m/z ##


spec_group <- 3

# Info for spectra group
xspectra[[spec_group]]

#Plotly allows for interactive check of m/z and intensity of ggplot2 objects
plotly::ggplotly(plot(xspectra[[spec_group]])) 


#plot EIC
pk_ids <- mcols(xspectra)$ms2_peak_id[[spec_group]]

rt_range <- chromPeaks(xdata)[pk_ids, c("rtmin", "rtmax")]
mz_range <- chromPeaks(xdata)[pk_ids, c("mzmin", "mzmax")]
pmz <- precursorMz(xspectra)[spec_group]
xdata_iwindow <- filterIsolationWindow(xdata, mz = pmz)
ms2_eics <- chromatogram(xdata_iwindow, rt = rt_range,
                         mz = mz_range, msLevel = 2L)

plot(ms2_eics)



## Currently need to do componentisation in order to group adducts and in source fragmentations



pdp <- PeakDensityParam(sampleGroups = sample.groups,
                          minFraction = minfrac, bw = bw.pre)
xdata <- groupChromPeaks(xdata, param = pdp)

pgp <- PeakGroupsParam(
    minFraction = 0.5)
xdata <- adjustRtime(xdata, param = pgp)

pdp <- PeakDensityParam(sampleGroups = sample.groups,
                          minFraction = minfrac, bw = bw.post)
xdata <- groupChromPeaks(xdata, param = pdp)

fpp <- FillChromPeaksParam(expandMz = 0, expandRt = 0, ppm = 0)
xdata <- fillChromPeaks(xdata, param = fpp)


```


