# A handbook on GC Q Exactive orbitrap use {#GCOrbitrapHandbook}



```{r msresolution, echo=FALSE}


MSresolution <- dplyr::tibble(
  Resolution_at_mz200 = c(7500, 15000, 30000, 45000, 60000, 120000, 240000),
  Scan_rate_Hz = c(40, 28, 12, 10, 7, 3, 1.5),
  Transient_lenght_ms = c(16, 32, 64, 96, 128, 256, 512),
  Suggested_max_fill_time_ms = c(11, 20, 50, 86, 110, 240, 500)
)
DT::datatable(MSresolution)


```



  

## Acquisition parameters

From: https://analyticalsciencejournals.onlinelibrary.wiley.com/doi/full/10.1002/rcm.8609?campaign=wolacceptedarticle
Orbitrap technology is based on ion oscillation frequency measurements around a central electrode, detected on a receiver plate.26 The mass spectrum is obtained after Fourier transformation of the oscillation signal with a two-point calibration. The longer the signal acquisition time, the better is the resolution.27 During this time, ions generated in the EI source are stored in an IT, the C-TRAP. The storage of ions in the C-TRAP plays a vital role in prematurely analysing discontinuous ions produced continuously by the chromatographic separation and the ionisation source. The ions introduced into the C-TRAP lose their energy in contact with N2 at approximately 1 mTorr.26 During an acquisition in full scan mode, it is possible to set the detector with different parameters: acquisition mass range, resolution, AGC target and tune values. The tune files gather the different settings of the mass spectrometer aiming to produce the ions of the EI source, but also the different voltages applied to the lenses in the ion path from the source to the C-TRAP. It is also possible to change the value of the C-TRAP energy offset (−5 to +5 V). For this study, we used standard conditions of EI (70 eV) in order to have the ideal conditions for reproducibility of the spectra. By default, the Orbitrap system acquires data in the range m/z 50–500 at 60 000 resolution mode and 70 eV for EI. The C-TRAP tune value was set to 0 V and AGT target to 1 × 106 ions with an automatic filling limit. 


- Recommended to have about 10-12 scans per peak.


  
```{r tuneparams, echo=FALSE}

# These are some optimum tune params


TuneParams <- dplyr::tibble(
  Parameter = c(
    "Repeller voltage (8-11 V)",
    "Source offset",
    "Lens 1",
    "Lens 2",
    "Lens 3",
    "Electron lens voltage",
    "Emission current (50-60 uA)",
    "IT (< 1)",
    "Normalized target level (NL) (clean source 1E9)",
    "Fore vacuum (< 6E-2 mbar)",
    "Ultrahigh vacuum (< 9E-10 mbar)"
  )

)
DT::datatable(TuneParams)

```


## Chemical ionization
If the instrument is in CI mode, choose the right Calibration gas and Ionization mode (“CI”), as well as the CI gas flow (typically 1-2), gas port (Port B as of 20210429, actual CI gas flow should indicate the correct port), and gas type (likely Methane, but check the cylinder). Open the CI gas flow at 2ml/min before opening anything else (before the calibration gas, filament, MS), otherwise the flow can be unstable. Afterwards the CI can be adjusted to lower flow 

  
## Terminologies
  
_Automatic Gain Control (AGC):_ Sets the ion injection time to maintain the optimum quantity of ions for each scan. With AGC on, the scan function consists of a prescan and an analytical scan. The split lens is used to start and stop the injection of ions into the mass analyzer. It provides a high deflection voltage most of the time so that ions are deflected into a baffle except when they are to be allowed into the C-Trap. The fast switching of the ion beam ensures the precise determination of the ion injection time that is required for AGC. This is since if too many ion are in the ion cloud going in to the orbitrap, then it will oversaturate the detector.
  
_Chemical Ionization (CI):_ 
  
_Nitrogen gas:_ Nitrogen gas is used for the purging of the ion source when removing the ion volume. It is also used in the C-Trap, when the ions are accumulated and their energy is dampened using a bath gas (nitrogen).



## XCMS preprocessing
Check: A scalable workflow to characterize the human exposome, SI-11, https://www.nature.com/articles/s41467-021-25840-9#MOESM14

IPO optimization


